Data Preparation

### 2021-07-13 x0_data-preparation.Rmd
my tweaks for the airport level

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(zoo)

@@

tmp %>%
  ggplot() +
  geom_line(aes(x = DATE, y = N), colour = "grey80") +
  geom_line(aes(x = DATE, y = N_sm, colour = THRES2, group = "RQ"), size = 2) +
  geom_line(aes(x = DATE, y = UP_THR), color = "blue") +
  geom_line(aes(x = DATE, y = LOW_THR), color = "red") +
  geom_line(aes(x = DATE, y = MED), color = "green") +
  geom_line(aes(x = DATE, y = BAND, group= "1"), colour = if_else(tmp$LOW_CRIT, "red","green"), size = 4) +
  geom_line(aes(x = DATE, y = BAND-500, group= "1"), colour = if_else(tmp$UP_CRIT, "red","green"), size = 4) +
  geom_line(aes(x = DATE, y = BAND-1500, group= "1", colour = STATE), size = 4) +
  theme_minimal() +
  theme(legend.position = "none")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(zoo)
library(patchwork)

@@

tmp %>%
  ggplot() +
  geom_line(aes(x = DATE, y = N), colour = "grey80") +
  geom_line(aes(x = DATE, y = N_sm, colour = THRES2, group = "RQ"), size = 2) +
  geom_line(aes(x = DATE, y = UP_THR), color = "blue") +
  geom_line(aes(x = DATE, y = LOW_THR), color = "red") +
  geom_line(aes(x = DATE, y = MED), color = "green") +
  geom_line(aes(x = DATE, y = BAND, group= "1"), colour = if_else(tmp$LOW_CRIT, "red","green"), size = 4) +
  geom_line(aes(x = DATE, y = BAND-500, group= "1"), colour = if_else(tmp$UP_CRIT, "red","green"), size = 4) +
  geom_line(aes(x = DATE, y = BAND-1500, group= "2", colour = STATE), size = 4) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Airport Level Considerations

top-5 in the whole sample, confirms preference :)

```{r}
full_dataset %>% 
  group_by(ADEP_REG, ADEP) %>% 
  summarise(N = n()) %>% 
  slice_max(order_by = N, n = 5)
```

```{r}
sbgr <- full_dataset %>% filter(ADEP == "SBGR")
```

```{r}
apt_timeline <- function(.df){
  df <- .df %>% group_by(DATE) %>%
    summarise(N_DEP = n()
              ,DEP_DOM = sum(ADEP_CTRY == ADES_CTRY)
              ,DEP_REG = sum(ADEP_REG == ADES_REG))
}

plot_timeline <- function(.df){
  df <- .df %>% 
  mutate(N_DEP_SM = roll_quantile(N_DEP, width = 7, p = .50))

  p <- df %>%  
  ggplot() +
  geom_line(aes(x = DATE, y = N_DEP), colour = "grey60") +
  geom_line(aes(x = DATE, y = N_DEP_SM)) +
  scale_x_date( limits = c(as.Date("2019-10-01"), as.Date("2021-06-01"))
               ,date_breaks = "60 days", date_labels = "%b\n%y") +
  theme_minimal()

  return(p)
}

sbgr %>% apt_timeline() %>% plot_timeline()
```

```{r}
sbgr_tl %>%
  ggplot() +
  geom_point(aes(x = DATE, y = DEP_DOM / N_DEP)) +
  theme_minimal()
```
```{r}
dep_vs_domshare <- function(.df){
  p <- .df %>%
  ggplot() +
  geom_point(aes(x = N_DEP, y = DEP_REG / N_DEP, colour = as.factor(year(DATE)))) +
  labs(colour = NULL) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)
                     ,expand = c(0,0)) +
  theme_minimal() +
  theme(legend.position = c(0.8, 0.8))
return(p)
}

sbgr_tl %>% dep_vs_domshare()
```

```{r}
check_apt <- function(.mvts, .apt){
  apt_mvts <- .mvts %>%
    filter(ADEP == .apt) %>% 
    apt_timeline()

  p1 <- apt_mvts  %>% plot_timeline()
  p2 <- apt_mvts %>% dep_vs_domshare()

  p <- (p1 + labs(subtitle = .apt)) + p2
return(p)
}

full_dataset %>% check_apt("EGLL")
```

```{r}
full_dataset %>% check_apt("EDDF")
```

```{r}
full_dataset %>% check_apt("LFPG")
```

```{r}
full_dataset %>% check_apt("KATL")
```

```{r}
full_dataset %>% check_apt("KORD")
```

```{r}
full_dataset %>% check_apt("KJFK")
```

```{r}
full_dataset %>% check_apt("EHAM")
```

```{r}
full_dataset %>% check_apt("EHAM")
```
